---
title: "MarkovChain_SPT_logseq"
author: "Katie Murenbeeld"
date: "6/7/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Intro:

Testing out markov chain for picking a "typical" sequence of logging events in Southern Idaho timber harvest activities.

```{r load_lib, echo=FALSE, eval=TRUE}
library(markovchain)
#library(dplyr)
library(tidyverse)
library(lubridate)
library(matrixStats)
```

## Load data

```{r data, echo=FALSE, eval=TRUE}

# Larger dataset for region
df_reg <- read.csv('proj_newcode3_20210607.csv')
df_reg$ACTION <- as.character(df_reg$ACTION)

# Smaller dataaset for S. Idaho
df_sid <- read.csv('proj_newcode3_20210607_v02.csv')
df_sid$ACTION <- as.character(df_sid$ACTION)

#df5$PROJ_DAYS <- ymd_hms(df5$PROJ_DAYS)

head(df_reg)
head(df_sid)

```


## Create probability(sequence) matrices

```{r seqmat, echo=TRUE, eval=TRUE}
# First dataset. I think too many action states. No binary Action - No Action?
seqmat_reg12 <- createSequenceMatrix(df_reg$ACTIVITY_TYPE, toRowProbs = TRUE, sanitize=TRUE)
seqmat_reg12

# New dataset with more Nat. Forests added. Very simplified activity types and has binary Action - No Action?
seqmat_reg2 <- createSequenceMatrix(df_reg$ACTION, toRowProbs = TRUE, sanitize=TRUE)
seqmat_reg2

seqmat_sid12 <- createSequenceMatrix(df_sid$ACTIVITY_TYPE, toRowProbs = TRUE, sanitize=TRUE)
seqmat_sid12

# Test datasets with only 2 projects from Nat. Forests in Idaho. Very simplified activity types and has binary Action - No Action?
seqmat_sid2 <- createSequenceMatrix(df_sid$ACTION, toRowProbs = TRUE, sanitize=TRUE)
seqmat_sid2

```

## Create the state vectors, need to be in alphabetical or ascending. Then create new markov chains.

```{r state_vec, echo=TRUE, eval=TRUE, fig.height=10, fig.width=15}

# I think I need the sequence vector in same order as a column or row of the transition matrix (alphabetical, from 0->1)?? 
# I think the new('markovchain',...) already checks the trans. matrix and makes sure the seq names and row/column names of the
# trans matrix correspond to each other.


seq_reg12 <- sort(unique(df_reg$ACTIVITY_TYPE))
seq_reg2 <- sort(unique(df_reg$ACTION))

seq_sid12 <- sort(unique(df_sid$ACTIVITY_TYPE))
seq_sid2 <- sort(unique(df_sid$ACTION))

markov_reg12 <- new('markovchain', 
               transitionMatrix = seqmat_reg12,
               states = seq_reg12)

markov_reg2 <- new('markovchain', 
               transitionMatrix = seqmat_reg2,
               states = seq_reg2)

markov_sid12 <- new('markovchain', 
               transitionMatrix = seqmat_sid12,
               states = seq_sid12)

markov_sid2 <- new('markovchain', 
               transitionMatrix = seqmat_sid2,
               states = seq_sid2)

plot(markov_reg12)
plot(markov_reg2)
plot(markov_sid12)
plot(markov_sid2)
```

## First Time Passage

```{r timepassage, echo=TRUE, eval=TRUE}
meanFirstPassageTime(markov_reg12)
meanFirstPassageTime(markov_reg2)
meanFirstPassageTime(markov_sid12)
meanFirstPassageTime(markov_sid2)
```

## From here I want to simulate a bunch of sequences and then look at a histogram for when project occur or not for each individual day.
## I also think I want to look at the mean and median project length in order to determine the n value for rmarkovchain. 

```{r mean_proj_dur, echo=TRUE, eval=TRUE}

projs_reg <- unique(unlist(df_reg$NEPA_PROJECT))

pd_df_reg = data.frame()

for (i in 1:length(projs_reg)) {
  #save project name to variable p
  p <- projs_reg[i]
  
  #create a temporary dataframe for each project
  tmp_df <- subset(df_reg, NEPA_PROJECT == projs_reg[i])
  
  # find the max project length for each project
  max_pd <- max(as.integer(tmp_df$PROJ_INT))
  
  pd_df_reg <- rbind(pd_df_reg, max_pd)
}
                  

projs_sid <- unique(unlist(df_sid$NEPA_PROJECT))

pd_df_sid = data.frame()

for (i in 1:length(projs_sid)) {
  #save project name to variable p
  p <- projs_sid[i]
  
  #create a temporary dataframe for each project
  tmp_df <- subset(df_sid, NEPA_PROJECT == projs_sid[i])
  
  # find the max project length for each project
  max_pd <- max(as.integer(tmp_df$PROJ_INT))
  
  pd_df_sid <- rbind(pd_df_sid, max_pd)
}
```

```{r hist_max_pd, echo=TRUE, eval=TRUE}

hist(pd_df_reg$X29L, breaks = 366)
hist(pd_df_sid$X1749L, breaks = 118)
mean(pd_df_reg$X29L)
mean(pd_df_sid$X1749L)
median(pd_df_reg$X29L)
median(pd_df_sid$X1749L)
```
## Let's replicate a bunch of times! Using the mean project durations

```{r, seq_sims_mean, echo=TRUE, eval=TRUE}
# mean project length

reg12_sims_mu <- replicate(20, {
  sequence <- rmarkovchain(n = 532, object = markov_reg12 , t0 = "NO_ACT")
})
  
reg2_sims_mu <- replicate(20, {
  sequence <- rmarkovchain(n = 532, object = markov_reg2, t0 = "0")
})

sid12_sims_mu <- replicate(20, {
  sequence <- rmarkovchain(n = 543, object = markov_sid12, t0 = "NO_ACT")
})

sid2_sims_mu <- replicate(20, {
  sequence <- rmarkovchain(n = 543, object = markov_sid2, t0 = "0")
})
```

## Let's replicate a bunch of times! Again! using the median project durations

```{r, seq_sims_med, echo=TRUE, eval=TRUE}
# mean project length

reg12_sims_med <- replicate(20, {
  sequence <- rmarkovchain(n = 149, object = markov_reg12 , t0 = "NO_ACT")
})
  
reg2_sims_med <- replicate(20, {
  sequence <- rmarkovchain(n = 149, object = markov_reg2, t0 = "0")
})

sid12_sims_med <- replicate(20, {
  sequence <- rmarkovchain(n = 190, object = markov_sid12, t0 = "NO_ACT")
})

sid2_sims_med <- replicate(20, {
  sequence <- rmarkovchain(n = 190, object = markov_sid2, t0 = "0")
})
```

## Starting from INT_CUT and 1

```{r, seq_sims_mean2, echo=TRUE, eval=TRUE}
# mean project length

reg12_sims_mu2 <- replicate(20, {
  sequence <- rmarkovchain(n = 532, object = markov_reg12 , t0 = "INT_CUT")
})
  
reg2_sims_mu2 <- replicate(20, {
  sequence <- rmarkovchain(n = 532, object = markov_reg2, t0 = "1")
})

sid12_sims_mu2 <- replicate(20, {
  sequence <- rmarkovchain(n = 543, object = markov_sid12, t0 = "INT_CUT")
})

sid2_sims_mu2 <- replicate(20, {
  sequence <- rmarkovchain(n = 543, object = markov_sid2, t0 = "1")
})
```

## Count the number of actions occuring on any specific day (row) of the synthetic sequence
## I think I need a for loop and also to transform the data frame? Look at EEB for examples.

```{r count_seq, echo=TRUE, eval=TRUE}
days <- c(1:532)


test_df <- as.data.frame(reg2_sims_mu2)
test_df$days <- c(1:532)

test_df2 <- t(as.matrix(test_df))
names(test_df2) <- c(1:532)
commmon_seqs <- colCounts(test_df2, value = "1")

barplot(commmon_seqs, names.arg = days)

```

## Well I forgot that my "numbers" 0,1 were actually characters.....which makes this a lot harder. Also, why so hard to loop over rows?



