---
title: "Forest_Sequencing"
author: "Katie Murenbeeld"
date: "4/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Intro:

Testing out markov chain for picking a "typical" sequence of logging events in Southern Idaho timber harvest activities. 

```{r load_lib, echo=FALSE, eval=TRUE}
library(markovchain)
#library(dplyr)
library(tidyverse)
```

## Load data

```{r data, echo=FALSE, eval=TRUE}
df <- read.csv('proj_newcode2_20210419.csv')
#df <- df %>% rename(df, SEQ_INT = int)
df$ACRES <- as.numeric(df$ACRES)

df1 <- read.csv('test_proj1_20210601.csv')
df1$ACT <- as.character(df1$ACT)

df2 <- read.csv('test_proj2_20210601.csv')
df2$ACT <- as.character(df2$ACT)

df3 <- read.csv('proj_newcode3_20210602.csv')
df3$ACTION <- as.character(df3$ACTION)

head(df)
head(df1)
head(df2)
head(df3)

df4 <- rbind(df1, df2)
head(df4)
```

## Try out the createSequenceMatrix fuction

```{r seqmat, echo=TRUE, eval=TRUE}
# First dataset. I think too many action states. No binary Action - No Action?
seqmat1 <- createSequenceMatrix(df$GEN_ACTIVITY, toRowProbs = TRUE, sanitize=TRUE)
seqmat1

# New dataset with more Nat. Forests added. Very simplified activity types and has binary Action - No Action?
seqmat2 <- createSequenceMatrix(df3$ACTIVITY_TYPE, toRowProbs = TRUE, sanitize=TRUE)
seqmat2

seqmat3 <- createSequenceMatrix(df3$ACTION, toRowProbs = TRUE, sanitize=TRUE)
seqmat3

# Test datasets with only 2 projects from Nat. Forests in Idaho. Very simplified activity types and has binary Action - No Action?
seqmat4 <- createSequenceMatrix(df4$ACT_TYPE, toRowProbs = TRUE, sanitize=TRUE)
seqmat4

seqmat5 <- createSequenceMatrix(df4$ACT, toRowProbs = TRUE, sanitize=TRUE)
seqmat5
```


```{r test_MCfit, echo=TRUE, eval=TRUE}
#mcFitMLE <- markovchainFit(data = df$GEN_ACTIVITY)
#summary(mcFitMLE)
```

```{r test_CTMC, echo=TRUE, eval=TRUE}
#df_ctmc <- df %>% select(GEN_ACTIVITY, SEQ_INT)

#ctmcFit(df_ctmc, byrow=TRUE, name="test_ctmc", confidencelevel = 0.95)
```

## This is fine, but I want to basically do this for each project? For each time an ACT_int is true? I guess I don't quite understand what the functions are doing.  I found a helpful tutorial to look into more: https://www.datacamp.com/community/tutorials/markov-chain-analysis-r as well as the cran documentation: https://cran.r-project.org/web/packages/markovchain/markovchain.pdf

```{r seq_verify, echo=TRUE, eval=TRUE}
seq1 <- unique(df$GEN_ACTIVITY)
seq1

verifyMarkovProperty(seq1)
```

```{r plot_test, echo=TRUE, eval=TRUE, fig.height=10, fig.width=15}
markov1 <- new('markovchain', 
               transitionMatrix = seqmat1,
               states = seq1)
layout <- matrix(c(-8,-8, 0,9, -8,0, 8,1, 0,-8, -8,8, 9,8, 8,-8, 0,0, -4,4, 4,4, 4,-4), ncol = 2, byrow = TRUE)
plot(markov1, edge.arrow.size=0.5, vertex.size=10, vertex.label.cex = 0.5, edge.label.cex=0.75,
     vertex.color="green", edge.arrow.mode = 3, layout=layout)
```


```{r plot_test2, echo=TRUE, eval=TRUE, fig.height=10, fig.width=15}

# I think I need the sequence vector in same order as a column or row of the transition matrix (alphabetical, from 0->1)?? 
# I think the new('markovchain',...) already checks the trans. matrix and makes sure the seq names and row/column names of the
# trans matrix correspond to each other.

seq2 <- unique(df3$ACTIVITY_TYPE)
seq2.1 <- sort(unique(df3$ACTIVITY_TYPE))

seq3 <- unique(df3$ACTION)
seq3.1 <- sort(unique(df3$ACTION))

markov2 <- new('markovchain', 
               transitionMatrix = seqmat2,
               states = seq2)

markov3 <- new('markovchain', 
               transitionMatrix = seqmat3,
               states = seq3)

markov2.1 <- new('markovchain', 
               transitionMatrix = seqmat2,
               states = seq2.1)

markov3.1 <- new('markovchain', 
               transitionMatrix = seqmat3,
               states = seq3.1)

plot(markov2)
plot(markov3)
plot(markov2.1)
plot(markov3.1)
```
```{r check_order, echo=TRUE, eval=TRUE}
# Check that the seq and transition matrix tables are in the same order.
print(seqmat2)
print(seq2)
print(seqmat3)
print(seq3)

print(seq2.1)
print(seq3.1)
```


```{r plot_test3, echo=TRUE, eval=TRUE, fig.height=10, fig.width=15}


#seqmat4 <- createSequenceMatrix(df2$ACT_TYPE, toRowProbs = TRUE, sanitize=TRUE)
#seqmat4 <- createSequenceMatrix(df2$ACT, toRowProbs = TRUE, sanitize=TRUE)

seq4 <- unique(df4$ACT_TYPE)
seq4.1 <- sort(unique(df4$ACT_TYPE))
seq5 <- unique(df4$ACT)
seq5.1 <- sort(unique(df4$ACT))
  

markov4 <- new('markovchain', 
               transitionMatrix = seqmat4,
               states = seq4)

markov5 <- new('markovchain', 
               transitionMatrix = seqmat5,
               states = seq5)

markov4.1 <- new('markovchain', 
               transitionMatrix = seqmat4,
               states = seq4.1)

markov5.1 <- new('markovchain', 
               transitionMatrix = seqmat5,
               states = seq5.1)

plot(markov4)
plot(markov5)
plot(markov4.1)
plot(markov5.1)
```

```{r preds7Days, echo=TRUE, eval=TRUE}
initialstate <- c(0,1)
initialstate2 <- c(0,1,0, 0,0,0, 0,0,0, 0,0,0)
initialstate3 <- c(0,0,0,1)

after7Days2 <- initialstate2 * (markov2^7) #large dataset, s = 12
after7Days3 <- initialstate * (markov3^7) #large dataset, s = 2
after7Days4 <- initialstate3 * (markov4^7) #2-project dataset, s = 4
after7Days5 <- initialstate * (markov5^7) #2-project dataset, s = 2

after7Days2
after7Days3
after7Days4
after7Days5
```

```{r preds30Days, echo=TRUE, eval=TRUE}
initialstate <- c(0,1)
initialstate2 <- c(0,1,0, 0,0,0, 0,0,0, 0,0,0)
initialstate3 <- c(0,0,0,1)

after30Days2 <- initialstate2 * (markov2^30) #large dataset, s = 12
after30Days3 <- initialstate * (markov3^30) #large dataset, s = 2
after30Days4 <- initialstate3 * (markov4^30) #2-project dataset, s = 4
after30Days5 <- initialstate * (markov5^30) #2-project dataset, s = 2

after30Days2
after30Days3
after30Days4
after30Days5
```

```{r plot_test4, echo=TRUE, eval=TRUE, fig.height=10, fig.width=15}
seqmat5 <- createSequenceMatrix(df3$ACT_TYPE, toRowProbs = TRUE, sanitize=TRUE)

seqmat6 <- createSequenceMatrix(df3$ACT, toRowProbs = TRUE, sanitize=TRUE)

seq5 <- unique(df3$ACT_TYPE)

seq6 <- unique(df3$ACT)
  

markov7 <- new('markovchain', 
               transitionMatrix = seqmat5,
               states = seq5)

markov8 <- new('markovchain', 
               transitionMatrix = seqmat6,
               states = seq6)
#layout <- matrix(c(-8,-8, 0,9, -8,0, 8,1, 0,-8, -8,8, 9,8, 8,-8, 0,0, -4,4, 4,4, 4,-4), ncol = 2, byrow = TRUE)
#plot(markov2, edge.arrow.size=0.5, vertex.size=10, vertex.label.cex = 0.5, edge.label.cex=0.75,
#     vertex.color="green", edge.arrow.mode = 3, layout=layout)

plot(markov7)
plot(markov8)
```

## I think I want to first, test out how to subset by project, and then try to do the subsetting in a loop. Somehow I want to get the transition matrices for each project. Still need to figure out how to get the time of the transitions. According to the datacamp markovchain in R "Transition matrices have the property that

```{r subset, echo=TRUE, eval=TRUE}
projects <- unique(df$NEPA_PROJECT)

proj1 <- subset(df, NEPA_PROJECT == projects[5])
#proj1

proj1_seq <- createSequenceMatrix(proj1$GEN_ACTIVITY, toRowProbs = TRUE, sanitize=TRUE)
proj1_seq
proj1
```
```{r loop_test, echo=TRUE, eval=TRUE}
projects <- unique(df$NEPA_PROJECT)

##for (x in projects) {
#  print(x)
#}

for (x in projects) {
  proj <- subset(df, NEPA_PROJECT == x)
  return(proj)
  proj_seq <- createSequenceMatrix(proj$SEQ_ACTIVITY, toRowProbs = TRUE, sanitize=TRUE)
  return(proj_seq)
}

proj_seq
```
```{r markov3, echo=TRUE, eval=TRUE}
markov3 <- new("markovchain", 
               states = c("EA-RH-FH", "INT_CUT"),
               transitionMatrix = proj1_seq,
               name = "test")
markov3
```

```{r mean_recur_time, echo=TRUE, eval=TRUE}
meantest <- meanRecurrenceTime(markov2)
meantest
```
```{r data_out_test, echo=TRUE, eval=TRUE}
mcDf <- as(markov3, "data.frame")
#mcNew <- as(mcDf, "markovchain")
mcDf
```

