---
title: "Forest_Sequencing"
author: "Katie Murenbeeld"
date: "4/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Intro:

Testing out markov chain for picking a "typical" sequence of logging events in Southern Idaho timber harvest activities. 

```{r load_lib, echo=FALSE, eval=TRUE}
library(markovchain)
#library(dplyr)
library(tidyverse)
```

## Load data

```{r data, echo=FALSE, eval=TRUE}
df <- read.csv('proj_newcode2_20210419.csv')
#df <- df %>% rename(df, SEQ_INT = int)
df$ACRES <- as.numeric(df$ACRES)

#df1 <- read.csv('test_proj1_20210601.csv')
#df1$ACT <- as.character(df1$ACT)

#df2 <- read.csv('test_proj2_20210601.csv')
#df2$ACT <- as.character(df2$ACT)

df3 <- read.csv('proj_newcode3_20210602.csv')
df3$ACTION <- as.character(df3$ACTION)

df5 <- read.csv('proj_newcode3_20210602_v02.csv')
df5$ACTION <- as.character(df5$ACTION)

df5$PROJ_DAYS <- as.integer(format(as.Date(df5$PROJ_DAYS), "%Y%m%d"))

head(df)
#head(df1)
#head(df2)
head(df3)
head(df5)
#df4 <- rbind(df1, df2)
#head(df4)
```

## Try out the createSequenceMatrix fuction

```{r seqmat, echo=TRUE, eval=TRUE}
# First dataset. I think too many action states. No binary Action - No Action?
seqmat1 <- createSequenceMatrix(df$GEN_ACTIVITY, toRowProbs = TRUE, sanitize=TRUE)
seqmat1

# New dataset with more Nat. Forests added. Very simplified activity types and has binary Action - No Action?
seqmat2 <- createSequenceMatrix(df3$ACTIVITY_TYPE, toRowProbs = TRUE, sanitize=TRUE)
seqmat2

seqmat3 <- createSequenceMatrix(df3$ACTION, toRowProbs = TRUE, sanitize=TRUE)
seqmat3

# Test datasets with only 2 projects from Nat. Forests in Idaho. Very simplified activity types and has binary Action - No Action?
seqmat4 <- createSequenceMatrix(df4$ACT_TYPE, toRowProbs = TRUE, sanitize=TRUE)
seqmat4

seqmat5 <- createSequenceMatrix(df4$ACT, toRowProbs = TRUE, sanitize=TRUE)
seqmat5


# New dataset with on SIdaho Nat.Forests. Very simplified activity types and has binary Action - No Action?
seqmat6 <- createSequenceMatrix(df5$ACTIVITY_TYPE, toRowProbs = TRUE, sanitize=TRUE)
seqmat6

seqmat7 <- createSequenceMatrix(df5$ACTION, toRowProbs = TRUE, sanitize=TRUE)
seqmat7
```


```{r test_MCfit, echo=TRUE, eval=TRUE}
#mcFitMLE <- markovchainFit(data = df$GEN_ACTIVITY)
#summary(mcFitMLE)
```

```{r test_CTMC, echo=TRUE, eval=TRUE}
#df_ctmc <- df %>% select(GEN_ACTIVITY, SEQ_INT)

#ctmcFit(df_ctmc, byrow=TRUE, name="test_ctmc", confidencelevel = 0.95)
```

## This is fine, but I want to basically do this for each project? For each time an ACT_int is true? I guess I don't quite understand what the functions are doing.  I found a helpful tutorial to look into more: https://www.datacamp.com/community/tutorials/markov-chain-analysis-r as well as the cran documentation: https://cran.r-project.org/web/packages/markovchain/markovchain.pdf

```{r seq_verify, echo=TRUE, eval=TRUE}
seq1 <- sort(unique(df$GEN_ACTIVITY))
seq1

verifyMarkovProperty(seq1)
```

```{r plot_test, echo=TRUE, eval=TRUE, fig.height=10, fig.width=15}
markov1 <- new('markovchain', 
               transitionMatrix = seqmat1,
               states = seq1)
layout <- matrix(c(-8,-8, 0,9, -8,0, 8,1, 0,-8, -8,8, 9,8, 8,-8, 0,0, -4,4, 4,4, 4,-4), ncol = 2, byrow = TRUE)
plot(markov1, edge.arrow.size=0.5, vertex.size=10, vertex.label.cex = 0.5, edge.label.cex=0.75,
     vertex.color="green", edge.arrow.mode = 3, layout=layout)
```


```{r plot_test2, echo=TRUE, eval=TRUE, fig.height=10, fig.width=15}

# I think I need the sequence vector in same order as a column or row of the transition matrix (alphabetical, from 0->1)?? 
# I think the new('markovchain',...) already checks the trans. matrix and makes sure the seq names and row/column names of the
# trans matrix correspond to each other.

seq2 <- unique(df3$ACTIVITY_TYPE)
seq2.1 <- sort(unique(df3$ACTIVITY_TYPE))

seq3 <- unique(df3$ACTION)
seq3.1 <- sort(unique(df3$ACTION))

markov2 <- new('markovchain', 
               transitionMatrix = seqmat2,
               states = seq2)

markov3 <- new('markovchain', 
               transitionMatrix = seqmat3,
               states = seq3)

markov2.1 <- new('markovchain', 
               transitionMatrix = seqmat2,
               states = seq2.1)

markov3.1 <- new('markovchain', 
               transitionMatrix = seqmat3,
               states = seq3.1)

plot(markov2)
plot(markov3)
plot(markov2.1)
plot(markov3.1)
```
```{r check_order, echo=TRUE, eval=TRUE}
# Check that the seq and transition matrix tables are in the same order.
print(seqmat2)
print(seq2)
print(seqmat3)
print(seq3)

print(seq2.1)
print(seq3.1)
```


```{r plot_test3, echo=TRUE, eval=TRUE, fig.height=10, fig.width=15}


#seqmat4 <- createSequenceMatrix(df2$ACT_TYPE, toRowProbs = TRUE, sanitize=TRUE)
#seqmat4 <- createSequenceMatrix(df2$ACT, toRowProbs = TRUE, sanitize=TRUE)

seq4 <- unique(df4$ACT_TYPE)
seq4.1 <- sort(unique(df4$ACT_TYPE))
seq5 <- unique(df4$ACT)
seq5.1 <- sort(unique(df4$ACT))
  

markov4 <- new('markovchain', 
               transitionMatrix = seqmat4,
               states = seq4)

markov5 <- new('markovchain', 
               transitionMatrix = seqmat5,
               states = seq5)

markov4.1 <- new('markovchain', 
               transitionMatrix = seqmat4,
               states = seq4.1)

markov5.1 <- new('markovchain', 
               transitionMatrix = seqmat5,
               states = seq5.1)

plot(markov4)
plot(markov5)
plot(markov4.1)
plot(markov5.1)
```
```{r plot_test4, echo=TRUE, eval=TRUE, fig.height=10, fig.width=15}
# For newer dataset with simplified activties, binary (action, no action). Only fo SIdaho National Forests.
seq6 <- unique(df5$ACTIVITY_TYPE)
seq6.1 <- sort(unique(df5$ACTIVITY_TYPE))
seq7 <- unique(df5$ACTION)
seq7.1 <- sort(unique(df5$ACTION))
  

markov6 <- new('markovchain', 
               transitionMatrix = seqmat6,
               states = seq6)

markov7 <- new('markovchain', 
               transitionMatrix = seqmat7,
               states = seq7)

markov6.1 <- new('markovchain', 
               transitionMatrix = seqmat6,
               states = seq6.1)

markov7.1 <- new('markovchain', 
               transitionMatrix = seqmat7,
               states = seq7.1)

plot(markov6)
plot(markov7)
plot(markov6.1)
plot(markov7.1)
```


```{r preds7Days, echo=TRUE, eval=TRUE}
initialstate <- c(0,1)
initialstate2 <- c(0,1,0, 0,0,0, 0,0,0, 0,0,0)
initialstate3 <- c(0,0,0,1)

after7Days2 <- initialstate2 * (markov2^7) #large dataset, s = 12
after7Days3 <- initialstate * (markov3^7) #large dataset, s = 2
after7Days4 <- initialstate3 * (markov4^7) #2-project dataset, s = 4
after7Days5 <- initialstate * (markov5^7) #2-project dataset, s = 2

after7Days2
after7Days3
after7Days4
after7Days5
```

```{r preds30Days, echo=TRUE, eval=TRUE}
initialstate <- c(0,1)
initialstate2 <- c(0,1,0, 0,0,0, 0,0,0, 0,0,0)
initialstate3 <- c(0,0,0,1)

after30Days2 <- initialstate2 * (markov2^30) #large dataset, s = 12
after30Days3 <- initialstate * (markov3^30) #large dataset, s = 2
after30Days4 <- initialstate3 * (markov4^30) #2-project dataset, s = 4
after30Days5 <- initialstate * (markov5^30) #2-project dataset, s = 2

after30Days2
after30Days3
after30Days4
after30Days5
```
```{r steady_states, echo=TRUE, eval=TRUE}

# order does matter for to find the steady state probability vector.
steadyStates(markov1)
steadyStates(markov2.1)
steadyStates(markov3.1)
steadyStates(markov4.1)
steadyStates(markov5.1)
```

```{r irreducible, echo=TRUE, eval=TRUE}

is.irreducible(markov1)
is.irreducible(markov2.1)
is.irreducible(markov3.1)
is.irreducible(markov4.1)
is.irreducible(markov5.1)

```

```{r period, echo=TRUE, eval=TRUE}

period(markov1)
period(markov2.1)
period(markov3.1)
period(markov4.1)
period(markov5.1)


```

```{r timepassage, echo=TRUE, eval=TRUE}
# Not sure I quite understand this yet...
firstPassagePdF <- firstPassage(object=markov3.1, state = "1", n=10)
#firstPassagePdF[1,1]


# I think this is what I want!
print("mean first passage time")
meanFirstPassageTime(markov3.1)
print("The expected number of days before a day when an activity occurs given that today no activity occurred is 131 days.")
print("The expected number of days before a day of no action given that today action occurred is 1.3 days.")
```

```{r timepassage2, echo=TRUE, eval=TRUE}
#Large dataset with Regional NFs.
# I think this is what I want!
print("mean first passage time")
meanFPT2 <- as.data.frame(meanFirstPassageTime(markov2.1))
meanFPT3 <- as.data.frame(meanFirstPassageTime(markov3.1))
#print("The expected number of days before a day when an activity occurs given that today no activity occurred is 131 days.")
#print("The expected number of days before a day of no action given that today action occurred is 1.3 days.")
```

```{r timepassage3, echo=TRUE, eval=TRUE}

# I think this is what I want! For smaller dataset with only 2 projects.
print("mean first passage time")
meanFirstPassageTime(markov4.1)
meanFirstPassageTime(markov5.1)
#print("The expected number of days before a day when an activity occurs given that today no activity occurred is 131 days.")
#print("The expected number of days before a day of no action given that today action occurred is 1.3 days.")
```

```{r timepassage4, echo=TRUE, eval=TRUE}
# With SIdaho NF dataset
# I think this is what I want!
print("mean first passage time")
meanFPT6 <- as.data.frame(meanFirstPassageTime(markov6.1))
meanFPT7 <- as.data.frame(meanFirstPassageTime(markov7.1))
#print("The expected number of days before a day when an activity occurs given that today no activity occurred is 131 days.")
#print("The expected number of days before a day of no action given that today action occurred is 1.3 days.")
```

```{r write_csv, echo=TRUE, eval=TRUE}
#write_csv(meanFPT2, file = "meanFPT2.csv")
#write_csv(meanFPT3, file = "meanFPT3.csv")
#write_csv(meanFPT6, file = "meanFPT6.csv")
#write_csv(meanFPT7, file = "meanFPT7.csv")
```


## Let's generate a random sequence according to the underlying stochastic process for the MCs I made above. 
It's simple. Just use rmarkovchain!

```{r rmark echo=TRUE, eval=TRUE}

# For 12 states
#markov6.1
sidaho_logseq1 <- rmarkovchain(n = 365, object = markov6.1 , t0 = "NO_ACT")
# For 2 states
sidaho_logseq2 <- rmarkovchain(n = 365, object = markov7.1, t0 = "0")

# For 12 states
reg_logseq1 <- rmarkovchain(n = 365, object = markov2.1, t0 = "NO_ACT")
# For 2 states
reg_logseq2 <- rmarkovchain(n = 365, object = markov3.1, t0 = "0")


```

```{r save_seq, echo=TRUE, eval=TRUE}
sidaho_logseq1 <- as.data.frame(sidaho_logseq1)
sidaho_logseq2 <- as.data.frame(sidaho_logseq2)
reg_logseq1 <- as.data.frame(reg_logseq1)
reg_logseq2 <- as.data.frame(reg_logseq2)

random_seg <- bind_cols(sidaho_logseq1, sidaho_logseq2, reg_logseq1, reg_logseq2)
```

## Let's replicate a bunch of times!

```{r, seq_sims, echo=TRUE, eval=TRUE}

sidaho_12states_sims <- replicate(20, {
  sequence <- rmarkovchain(n = 365, object = markov6.1 , t0 = "NO_ACT")
})
  
sidaho_2states_sims <- replicate(20, {
  sequence <- rmarkovchain(n = 365, object = markov7.1, t0 = "0")
})

reg_12states_sims <- replicate(20, {
  sequence <- rmarkovchain(n = 365, object = markov2.1, t0 = "NO_ACT")
})

reg_2states_sims <- replicate(20, {
  sequence <- rmarkovchain(n = 365, object = markov3.1, t0 = "0")
})
```

```{r save_seq2, echo=TRUE, eval=TRUE}

```


## From here I want to simulate a bunch of sequences and then look at a histogram for when project occur or not for each individual day.
## I also think I want to look at the mean and median project length in order to determine the n value for rmarkovchain. 

```{r mean_proj_dur, echo=TRUE, eval=TRUE}

projs3 <- unique(unlist(df3$NEPA_PROJECT))

pd_df3 = data.frame()

for (i in 1:length(projs3)) {
  #save project name to variable p
  p <- projs3[i]
  
  #create a temporary dataframe for each project
  tmp_df <- subset(df3, NEPA_PROJECT == projs3[i])
  
  # find the max project length for each project
  max_pd <- max(as.integer(tmp_df$PROJ_DAYS))
  
  pd_df3 <- rbind(pd_df3, max_pd)
}
                  

projs5 <- unique(unlist(df5$NEPA_PROJECT))


```













## I think I want to first, test out how to subset by project, and then try to do the subsetting in a loop. Somehow I want to get the transition matrices for each project. Still need to figure out how to get the time of the transitions. According to the datacamp markovchain in R "Transition matrices have the property that

```{r subset, echo=TRUE, eval=TRUE}
projects <- unique(df$NEPA_PROJECT)

proj1 <- subset(df, NEPA_PROJECT == projects[5])
#proj1

proj1_seq <- createSequenceMatrix(proj1$GEN_ACTIVITY, toRowProbs = TRUE, sanitize=TRUE)
proj1_seq
proj1
```
```{r loop_test, echo=TRUE, eval=TRUE}
projects <- unique(df$NEPA_PROJECT)

##for (x in projects) {
#  print(x)
#}

for (x in projects) {
  proj <- subset(df, NEPA_PROJECT == x)
  return(proj)
  proj_seq <- createSequenceMatrix(proj$SEQ_ACTIVITY, toRowProbs = TRUE, sanitize=TRUE)
  return(proj_seq)
}

proj_seq
```
```{r markov3, echo=TRUE, eval=TRUE}
markov3 <- new("markovchain", 
               states = c("EA-RH-FH", "INT_CUT"),
               transitionMatrix = proj1_seq,
               name = "test")
markov3
```

```{r mean_recur_time, echo=TRUE, eval=TRUE}
meantest <- meanRecurrenceTime(markov2)
meantest
```
```{r data_out_test, echo=TRUE, eval=TRUE}
mcDf <- as(markov3, "data.frame")
#mcNew <- as(mcDf, "markovchain")
mcDf
```

